Multi-stage programming with functors and monads: eliminating abstraction overhead from generic code

### 2008-06-23 SCP2008/paper/metamonads.tex
include .bbl directly into document for submission

% Embed .bbl file directly to make document self-contained
\bibliography{metamonads}
\bibliographystyle{elsart-num}

%% \begin{thebibliography}{10}
%% \expandafter\ifx\csname url\endcsname\relax
%%   \def\url#1{\texttt{#1}}\fi
%% \expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi

>>>>

% \bibliography{metamonads}
% \bibliographystyle{elsart-num}

% Embed .bbl file directly to make document self-contained
\begin{thebibliography}{10}
\expandafter\ifx\csname url\endcsname\relax
  \def\url#1{\texttt{#1}}\fi
\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi

\bibitem{Carette06}
J.~Carette, Gaussian {E}limination: a case study in efficient genericity with
  {MetaOCaml}, Science of Computer Programming 62~(1) (2006) 3--24, special
  Issue on the First MetaOCaml Workshop 2004.

\bibitem{Monagan:2001:M7P}
M.~B. Monagan, K.~O. Geddes, K.~M. Heal, G.~Labahn, S.~M. Vorkoetter,
  J.~McCarron, P.~DeMarco, Maple 7 Programming Guide, Waterloo Maple Inc.,
  2001.

\bibitem{Gruntz:1994:IG}
D.~Gruntz, M.~Monagan, Introduction to {Gauss}, \textsc{SIGSAM Bulletin}:
  Communications on Computer Algebra 28~(2) (1994) 3--19.

\bibitem{Axiom}
R.~D. Jenks, R.~S. Sutor, {AXIOM}: The Scientific Computation System, Springer
  Verlag, 1992.

\bibitem{Kennedy01Telescoping}
K.~Kennedy, B.~Broom, K.~Cooper, J.~Dongarra, R.~Fowler, D.~Gannon,
  L.~Johnsson, J.~Mellor-Crummey, L.~Torczon, Telescoping languages: A strategy
  for automatic generation of scientific problem-solving systems from annotated
  libraries, Journal of Parallel and Distributed Computing 61~(12) (2001)
  1803--1826.

\bibitem{Veldhuizen:2004}
T.~L. Veldhuizen, Active libraries and universal languages, Ph.D. thesis,
  Indiana University Computer Science (May 2004).
\newline\urlprefix\url{http://osl.iu.edu/~tveldhui/papers/2004/dissertation.pd%
f}

\bibitem{scp-CohenDGHKP06}
A.~Cohen, S.~Donadio, M.~J. Garzar{\'a}n, C.~A. Herrmann, O.~Kiselyov, D.~A.
  Padua, In search of a program generator to implement generic transformations
  for high-performance computing, Sci. {C}omput. {P}rogram 62~(1) (2006)
  25--46.

\bibitem{Czarnecki}
K.~Czarnecki, U.~W. Eisenecker, Generative programming: methods, tools, and
  applications, ACM Press/Addison-Wesley Publishing Co., 2000.

\bibitem{Veldhuizen:1998:ISCOPE}
T.~L. Veldhuizen, Arrays in {Blitz++}, in: Proceedings of the 2nd International
  Scientific Computing in Object-Oriented Parallel Environments (ISCOPE'98),
  Lecture Notes in Computer Science, Springer-Verlag, 1998, pp. 223--230.

\bibitem{musser89generic}
D.~R. Musser, A.~A. Stepanov, Generic programming, in: {ISSAC 1988}:
  Proceedings of the International Symposium on Symbolic and Algebraic
  Computation, Vol. 358 of Lecture Notes in Computer Science, Springer-Verlag,
  1989, pp. 13--25.

\bibitem{musser94algorithmoriented}
D.~R. Musser, A.~A. Stepanov, Algorithm-oriented generic libraries, Software -
  Practice and Experience 24~(7) (1994) 623--642.

\bibitem{BOOST}
J.~Siek, L.-Q. Lee, A.~Lumsdaine, The Boost Graph Library: User Guide and
  Reference Manual, Addison-Wesley, 2002.

\bibitem{POOMA}
I.~John V.W.~Reynders, J.~C. Cummings, The {POOMA} framework, Comput. Phys.
  12~(5) (1998) 453--459.

\bibitem{ATLAS}
R.~C. Whaley, A.~Petitet, J.~J. Dongarra, Automated empirical optimization of
  software and the {ATLAS} project, Parallel Computing 27~(1--2) (2001) 3--35.

\bibitem{HygienicMacros}
E.~E. Kohlbecker, D.~P. Friedman, M.~Felleisen, B.~F. Duba, Hygienic macro
  expansion, in: LISP and Functional Programming, 1986, pp. 151--161.

\bibitem{camlp4}
D.~de~Rauglaudre, Camlp4 reference manual,
  \url{http://caml.inria.fr/camlp4/manual/} (Jan. 2002).

\bibitem{conf/dagstuhl/CzarneckiOST03}
K.~Czarnecki, J.~T. O'Donnell, J.~Striegnitz, W.~Taha, {DSL} implementation in
  {MetaOCaml}, {Template Haskell}, and {C++}, in: C.~Lengauer, D.~S. Batory,
  C.~Consel, M.~Odersky (Eds.), Domain-Specific Program Generation, Vol. 3016
  of Lecture Notes in Computer Science, Springer, 2003, pp. 51--72.

\bibitem{CTHL03}
C.~Calcagno, W.~Taha, L.~Huang, X.~Leroy, Implementing multi-stage languages
  using asts, gensym, and reflection, in: F.~Pfenning, Y.~Smaragdakis (Eds.),
  GPCE, Vol. 2830 of Lecture Notes in Computer Science, Springer, 2003, pp.
  57--76.

\bibitem{metaocaml-org}
{MetaOCaml}, \url{http://www.metaocaml.org}.

\bibitem{TahaSheard97}
W.~Taha, T.~Sheard, Multi-stage programming with explicit annotations, in:
  Proceedings of the Symposium on Partial Evaluation and Semantic-Based Program
  Manipulation {(PEPM)}, ACM, Amsterdam, 1997, pp. 203--217.

\bibitem{TahaThesis}
W.~Taha, Multi-stage programming: its theory and applications, Ph.D. thesis,
  Oregon Graduate Institute of Science and Technology (1999).

\bibitem{Taha2000}
W.~Taha, A sound reduction semantics for untyped {CBN} multi-stage computation.
  {Or}, the theory of {MetaML} is non-trival, in: PEPM, 2000, pp. 34--43.

\bibitem{Pueschel:05}
M.~P{\"u}schel, J.~M.~F. Moura, J.~Johnson, D.~Padua, M.~Veloso, B.~W. Singer,
  J.~Xiong, F.~Franchetti, A.~Ga\v{c}i\'{c}, Y.~Voronenko, K.~Chen, R.~W.
  Johnson, N.~Rizzolo, {SPIRAL}: Code generation for {DSP} transforms,
  Proceedings of the IEEE, special issue on "Program Generation, Optimization,
  and Adaptation" 93~(2).

\bibitem{dongarra7}
Z.~Chen, J.~Dongarra, P.~Luszczek, K.~Rothe, Lapack for clusters project: An
  example of self adapting numerical software, Hawaii International Conference
  on System Sciences HICSS-37.

\bibitem{KiselyovTaha}
O.~Kiselyov, K.~N. Swadi, W.~Taha, A methodology for generating verified
  combinatorial circuits, in: EMSOFT '04: Proceedings of the fourth ACM
  international conference on Embedded software, ACM, 2004, pp. 249--258.

\bibitem{CaretteKiselyov05}
J.~Carette, O.~Kiselyov, Multi-stage programming with {F}unctors and {M}onads:
  Eliminating abstraction overhead from generic code, in: Gl{\"u}ck and Lowry
  \cite{DBLP:conf/gpce/2005}, pp. 256--274.

\bibitem{SwadiMonadic06}
K.~N. Swadi, W.~Taha, O.~Kiselyov, E.~Pasalic, A monadic approach for avoiding
  code duplication when staging memoized functions, in: J.~Hatcliff, F.~Tip
  (Eds.), PEPM, ACM, 2006, pp. 160--169.

\bibitem{metamonadsURL}
{Source code}, \url{http://www.cas.mcmaster.ca/~carette/metamonads/}.

\bibitem{Golub-vanLoan}
G.~H. {Golub}, C.~F. {Van Loan}, Matrix Computations, 3rd Edition, Johns
  Hopkins University Press, Baltimore, MD, 1996.

\bibitem{kiczales97aspectoriented}
G.~Kiczales, J.~Lamping, A.~Menhdhekar, C.~Maeda, C.~Lopes, J.-M. Loingtier,
  J.~Irwin, Aspect-oriented programming, in: M.~Ak\c{s}it, S.~Matsuoka (Eds.),
  Proceedings European Conference on Object-Oriented Programming, Vol. 1241,
  Springer-Verlag, 1997, pp. 220--242.

\bibitem{709568}
J.~Irwin, J.-M. Loingtier, J.~R. Gilbert, G.~Kiczales, J.~Lamping,
  A.~Mendhekar, T.~Shpeisman, Aspect-oriented programming of sparse matrix
  code, in: ISCOPE '97: Proceedings of the Scientific Computing in
  Object-Oriented Parallel Environments, Springer-Verlag, London, UK, 1997, pp.
  249--256.

\bibitem{mendhekar97rg}
A.~Mendhekar, G.~Kiczales, J.~Lamping, {RG}: {A} case-study for aspect-oriented
  programming, Tech. Rep. SPL97-009 P9710044, {X}erox {PARC}, Palo Alto, CA,
  USA (February 1997).

\bibitem{journals/cacm/parnas72a}
D.~L. Parnas, On the criteria to be used in decomposing systems into modules,
  Commun. ACM 15~(12) (1972) 1053--1058.

\bibitem{journals/tse/Parnas76}
D.~L. Parnas, On the design and development of program families, IEEE Trans.
  Software Eng. 2~(1) (1976) 1--9.

\bibitem{EWD:EWD447}
E.~W. Dijkstra, On the role of scientific thought, published as
  \cite{EWD:EWD447pub}.
\newline\urlprefix\url{http://www.cs.utexas.edu/users/EWD/ewd04xx/EWD447.PDF}

\bibitem{ZhCaJeMo06a}
W.~Zhou, J.~Carette, D.~J. Jeffrey, M.~B. Monagan, Hierarchical representations
  with signatures for large expression management, in: Proceedings of
  Artificial Intelligence and Symbolic Computation, Vol. 4120 of Lecture Notes
  in Computer Science, 2006, pp. 254--268.

\bibitem{Bondorf:92}
A.~Bondorf, Improving binding times without explicit {CPS}-conversion, in: 1992
  ACM Conference on Lisp and Functional Programming. San Francisco, California,
  1992, pp. 1--10.

\bibitem{conf/pepm/BondorfD94}
A.~Bondorf, D.~Dussart, Improving {CPS-}based partial evaluation: Writing cogen
  by hand, in: {ACM} {SIGPLAN} Workshop on Partial Evaluation and
  Semantics-Based Program Manipulation, Orlando, Florida, 1994, pp. 1--9.

\bibitem{Haskell98Report}
S.~Peyton~Jones, et~al., The Revised Haskell 98 Report, Cambridge Univ.\null{}
  Press, 2003, also on \url{http://haskell.org/}.

\bibitem{moggi-notions}
E.~Moggi, Notions of computation and monads, Information and Computation 93~(1)
  (1991) 55--92.

\bibitem{Filinski:Representing}
A.~Filinski, Representing monads, in: POPL, 1994, pp. 446--457.

\bibitem{liang-interpreter}
S.~Liang, P.~Hudak, M.~Jones, Monad transformers and modular interpreters, in:
  {POPL} '95: Conference Record of the Annual {ACM} Symposium on Principles of
  Programming Languages, {ACM} {P}ress, {N}ew {Y}ork, 1995, pp. 333--343.

\bibitem{taha-environment}
W.~Taha, M.~F. Nielsen, Environment classifiers, in: {POPL} '03: Conference
  Record of the Annual {ACM} Symposium on Principles of Programming Languages,
  {ACM} {P}ress, {N}ew {Y}ork, 2003, pp. 26--37.

\bibitem{mlton-proplist}
{{MLton} team}, Property list, \url{http://mlton.org/PropertyList}.

\bibitem{BurdenFaires89}
R.~L. Burden, J.~D. Faires, Numerical analysis: 4th ed, PWS Publishing Co.,
  Boston, MA, USA, 1989.

\bibitem{Bareiss68}
E.~H. Bareiss, {S}ylvester's identity and multistep {G}aussian {Elimination},
  Math. Comput. 22~(103) (1968) 565--579.

\bibitem{offshoring}
J.~Eckhardt, R.~Kaiabachev, E.~Pasalic, K.~N. Swadi, W.~Taha, Implicitly
  heterogeneous multi-stage programming, in: Gl{\"u}ck and Lowry
  \cite{DBLP:conf/gpce/2005}, pp. 275--292.

\bibitem{Watt:2002:HCA}
S.~M. Watt, Aldor, in: J.~Grabmeier, E.~Kaltofen, V.~Weispfennig (Eds.),
  Computer Algebra Handbook: Foundations, Applications, Systems, Springer
  Verlag, 2003, pp. 265--270.

\bibitem{Gluck95}
R.~Gl{\"u}ck, R.~Nakashige, R.~Z{\"o}chling, Binding-time analysis applied to
  mathematical algorithms, in: System Modelling and Optimization, Chapman and
  Hall, 1995, pp. 137--146.

\bibitem{GluckJ97}
R.~Gl{\"u}ck, J.~J{\o}rgensen, An automatic program generator for multi-level
  specialization, Lisp and Symbolic Computation 10~(2) (1997) 113--158.

\bibitem{edsl}
P.~Hudak, Building domain-specific embedded languages, ACM Comput. Surv.
  28~(4es) (1996) 196.

\bibitem{WhenHowDSL}
M.~Mernik, J.~Heering, A.~M. Sloane, When and how to develop domain-specific
  languages, ACM Comput. Surv. 37~(4) (2005) 316--344.

\bibitem{DBLP:conf/gpce/2005}
R.~Gl{\"u}ck, M.~R. Lowry (Eds.), Generative Programming and Component
  Engineering, 4th Internation Conference, GPCE 2005, Tallinn, Estonia,
  September 29 - October 1, 2005, Proceedings, Vol. 3676 of Lecture Notes in
  Computer Science, Springer, 2005.

\bibitem{EWD:EWD447pub}
E.~W. Dijkstra, On the role of scientific thought, in: Selected Writings on
  Computing: A Personal Perspective, Springer-Verlag, 1982, pp. 60--66.

\end{thebibliography}
